---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mfair: Matrix Factorization with Auxiliary Information in R

<!-- badges: start -->
<!-- badges: end -->

Methods for matrix factorization to leverage auxiliary information based on the paper [MFAI: A scalable Bayesian matrix factorization approach to leveraging auxiliary information](https://doi.org/10.48550/arXiv.2303.02566).
MFAI integrates gradient boosted trees in the probabilistic matrix factorization framework to effectively leverage auxiliary information.
The parameters in MAFI can be automatically determined under the empirical Bayes framework, making it adaptive to the utilization of auxiliary information and immune to overfitting.

## Installation

You can install the development version of `mfair` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("YangLabHKUST/mfair")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example1}
set.seed(20230306)
library(mfair)

# Simulate data
# Set the data dimension and rank
N <- 100
M <- 100
K_true <- 2L

# Set the proportion of variance explained (PVE)
PVE_Z <- 0.8
PVE_Y <- 0.5

# Generate auxiliary information X
X1 <- runif(N, min = -10, max = 10)
X2 <- runif(N, min = -10, max = 10)
X <- cbind(X1, X2)

# F(X)
FX1 <- X1 / 2 - X2
FX2 <- (X1^2 - X2^2 + 2 * X1 * X2) / 10
FX <- cbind(FX1, FX2)

# Generate loadings Z (= F(X) + noise)
sig1_sq <- var(FX1) * (1 / PVE_Z - 1)
Z1 <- FX1 + rnorm(n = N, mean = 0, sd = sqrt(sig1_sq))
sig2_sq <- var(FX2) * (1 / PVE_Z - 1)
Z2 <- FX2 + rnorm(n = N, mean = 0, sd = sqrt(sig2_sq))
Z <- cbind(Z1, Z2)

# Generate factors W
W <- matrix(rnorm(M * K_true), nrow = M, ncol = K_true)

# Generate the main data matrix Y_obs (= Y + noise)
Y <- Z %*% t(W)
Y_var <- var(as.vector(Y))
epsilon_sq <- Y_var * (1 / PVE_Y - 1)
Y_obs <- Y + matrix(
  rnorm(N * M,
    mean = 0,
    sd = sqrt(epsilon_sq)
  ),
  nrow = N, ncol = M
)

# Create MFAIR object
mfairObject <- createMFAIR(Y_obs, X, K_max = K_true)

# Fit the MFAI model
mfairObject <- fitGreedy(mfairObject, sf_para = list(verbose_loop = FALSE))

# Prediction based on the low-rank approximation
Y_hat <- predict(mfairObject)

# Root-mean-square-error
sqrt(mean((Y_obs - Y_hat)^2))

# Predicted/true matrix variance ratio
var(as.vector(Y_hat)) / var(as.vector(Y_obs))

# Prediction/noise variance ratio
var(as.vector(Y_hat)) / var(as.vector(Y_obs - Y_hat))
```

`mfair` can also handle the matrix with missing entries:
```{r example2}
# Split the data into the training set and test set
n_all <- N * M
training_ratio <- 0.5
train_set <- sample(1:n_all, n_all * training_ratio, replace = FALSE)
Y_train <- Y_test <- Y_obs
Y_train[-train_set] <- NA
Y_test[train_set] <- NA

# Create MFAIR object
mfairObject <- createMFAIR(Y_train, X, K_max = K_true)

# Fit the MFAI model
mfairObject <- fitGreedy(mfairObject, sf_para = list(verbose_loop = FALSE))

# Prediction based on the low-rank approximation
Y_hat <- predict(mfairObject)

# Root-mean-square-error
sqrt(mean((Y_test - Y_hat)^2, na.rm = TRUE))

# Predicted/true matrix variance ratio
var(as.vector(Y_hat), na.rm = TRUE) / var(as.vector(Y_obs), na.rm = TRUE)

# Prediction/noise variance ratio
var(as.vector(Y_hat), na.rm = TRUE) / var(as.vector(Y_obs - Y_hat), na.rm = TRUE)
```

For more documentation and examples, please visit our package [website](https://yanglabhkust.github.io/mfair/).


## Citing our work

If you find the `mfair` package or any of the source code in this repository useful for your work, please cite:

> Wang, Z., Zhang, F., Zheng, C., Hu, X., Cai, M., and Yang, C. (2023). MFAI: A scalable Bayesian matrix factorization approach to leveraging auxiliary information. arXiv preprint arXiv:2303.02566. URL: https://doi.org/10.48550/arXiv.2303.02566.

## Development

The package is developed by Zhiwei Wang (<zhiwei.wang@connect.ust.hk>).

## Contact Information

Please feel free to contact Zhiwei Wang (<zhiwei.wang@connect.ust.hk>), Prof. Mingxuan Cai (<mingxcai@cityu.edu.hk>), or Prof. Can Yang (<macyang@ust.hk>) if any inquiries.
