---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mfair: Matrix Factorization with Auxiliary Information in R

<!-- badges: start -->
<!-- badges: end -->

Methods for matrix factorization to leverage auxiliary information based on the paper MFAI.
The name of the package, `mfair`, comes from "Matrix Factorization with Auxiliary Information in R".

## Installation

You can install the development version of `mfair` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("YangLabHKUST/mfair")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(mfair)
## basic example code
```

## Development

The package `mfair` is developed by Zhiwei Wang (<zhiwei.wang@connect.ust.hk>).

## Contact Information

Please feel free to contact Zhiwei Wang (<zhiwei.wang@connect.ust.hk>), Prof. Mingxuan Cai (<mingxcai@cityu.edu.hk>), or Prof. Can Yang (<macyang@ust.hk>) with any inquiries.
