---
title: "Enrichment from movie genre information"
author: "WANG Zhiwei, HKUST"
date: "May 14, 2023"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Enrichment from movie genre information}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(RhpcBLASctl)
# blas_set_num_threads(32)
# install.packages("devtools")
# devtools::install_github("YangLabHKUST/mfair")
library(mfair)
library(reshape2)
library(ggplot2)
library(scales)
set.seed(20230514)
```

# Fitting the MFAI model

We first set the `rating` matrix as the main data matrix $Y$, and the `genre` matrix as the auxiliary matrix $X$.
We then proceed to fit the MFAI model.

```{r}
# Create MFAIR object
Y <- t(ml100k$rating)
X <- ml100k$genre
mfairObject <- createMFAIR(Y, X, K_max = 3)

# Fit the MFAI model
mfairObject <- fitGreedy(mfairObject,
  save_init = TRUE,
  sf_para = list(verbose_loop = FALSE)
)
```

# Importance score

After fitting the MFAI model, we can use the `getImportance()` function to obtain the importance score of each genre within each factor.

```{r}
# Show top three factors
n_factors <- 3
importance <- as.data.frame(getImportance(mfairObject, which_factors = seq_len(n_factors)))
importance$Genre <- rownames(importance)
importance_long <- melt(
  data = importance,
  id.vars = "Genre",
  variable.name = "Factor",
  value.name = "Importance"
)
importance_long$Genre <- factor(importance_long$Genre, levels = rev(colnames(X)))
# head(importance_long)
```

```{r}
# Visualize the importance score
p1 <- ggplot(
  data = importance_long,
  aes(x = Genre, y = Importance, fill = Genre)
) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = label_comma(accuracy = 1)) +
  xlab(NULL) +
  ylab("Importance score") +
  guides(fill = "none") +
  theme_bw() +
  theme(
    text = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    aspect.ratio = 2
  ) +
  facet_grid(. ~ Factor, scales = "free")
p1
```

The higher the importance score is, the more a specific movie genre contributes to improving the model.

# Negative control

Next, let's create a permuted movie genre matrix $X^{\text{pmt}}$, where each column of $X^{\text{pmt}}$ was obtained by permuting the entries of the corresponding column in the real genre matrix $X$.
Then we fit the MFAI model with $Y$ and $X^{\text{pmt}}$ as input.

```{r}
n_pmt <- dim(X)[2]

X_pmt <- apply(X,
  MARGIN = 2,
  FUN = function(x) {
    N <- length(x)
    x[sample(x = c(1:N), size = N, replace = FALSE)]
  }
)

colnames(X_pmt) <- paste0(colnames(X), "_permuted")
```

```{r}
# Create MFAIR object and use the same initialization
mfairObject_pmt <- createMFAIR(Y, X_pmt, K_max = 3)
mfairObject_pmt@initialization <- mfairObject@initialization

# Fit the MFAI model
mfairObject_pmt <- fitGreedy(mfairObject_pmt,
  sf_para = list(verbose_loop = FALSE)
)
```

```{r}
# Show top three factors
n_factors <- 3
importance_pmt <- as.data.frame(getImportance(mfairObject_pmt, which_factors = seq_len(n_factors)))
importance_pmt$Genre <- rownames(importance_pmt)
importance_pmt_long <- melt(
  data = importance_pmt,
  id.vars = "Genre",
  variable.name = "Factor",
  value.name = "Importance"
)
importance_pmt_long$Genre <- factor(importance_pmt_long$Genre, levels = rev(colnames(X_pmt)))
# head(importance_pmt_long)

# Visualize the importance score
p2 <- ggplot(
  data = importance_pmt_long,
  aes(x = Genre, y = Importance, fill = Genre)
) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = label_comma(accuracy = 1)) +
  xlab(NULL) +
  ylab("Importance score") +
  guides(fill = "none") +
  theme_bw() +
  theme(
    text = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    aspect.ratio = 2
  ) +
  facet_grid(. ~ Factor, scales = "free")
p2
```

MFAI correctly assigns low importance scores to all permuted genres, suggesting that MFAI avoids incorporating irrelevant auxiliary information. 

At last, we use $X^{\text{both}} = [X, X^{\text{pmt}}]$ as the input auxiliary matrix and fit the MFAI model.

```{r}
X_both <- cbind(X, X_pmt)

# Create MFAIR object and use the same initialization
mfairObject_both <- createMFAIR(Y, X_both, K_max = 3)
mfairObject_both@initialization <- mfairObject@initialization

# Fit the MFAI model
mfairObject_both <- fitGreedy(mfairObject_both,
  sf_para = list(verbose_loop = FALSE)
)
```

```{r}
# Show top three factors
n_factors <- 3
importance_both <- as.data.frame(getImportance(mfairObject_both, which_factors = seq_len(n_factors)))
importance_both$Genre <- rownames(importance_both)
importance_both_long <- melt(
  data = importance_both,
  id.vars = "Genre",
  variable.name = "Factor",
  value.name = "Importance"
)
importance_both_long$Genre <- factor(importance_both_long$Genre, levels = rev(colnames(X_both)))
# head(importance_both_long)

# Visualize the importance score
p3 <- ggplot(
  data = importance_both_long,
  aes(x = Genre, y = Importance, fill = Genre)
) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = label_comma(accuracy = 1)) +
  xlab(NULL) +
  ylab("Importance score") +
  guides(fill = "none") +
  theme_bw() +
  theme(
    text = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 8),
    aspect.ratio = 2
  ) +
  facet_grid(. ~ Factor, scales = "free")
p3
```

MFAI successfully distinguished the useful movie genres from irrelevant ones.
Moreover, the importance scores obtained using $X^{\text{both}}$ are highly consistent with those obtained using $X$ and $X^{\text{pmt}}$ as separate inputs, indicating the stability and robustness of MFAI.
